Towards a classification framework for social machines

### 2013-04-04 latex/paper_new.tex
Updated to version submitted to sheridan

\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Nigel Shadbolt,^{1} Daniel Alexander Smith,^{2} Elena Simperl,^{3} Max Van Kleek,^{4} Yang Yang,^{5} Wendy Hall^{6}
       \affaddr{Web and Internet Science Group}\\
       \affaddr{University of Southampton}\\
       \affaddr{Southampton, UK}\\
       \email{\{^{1}nrs, ^{2}ds, ^{3}es, ^{4}emax, ^{5}yy, ^{6}wh@ecs.soton.ac.uk}
}

% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

The remaining sections are organised as follows. In order to pin down what constitutes a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} provides a summary of this comparative analysis. In Section \ref{sec:framework} we present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via repertory grid analysis. We conclude with our proposed future work, including details concerning the evaluation of the framework and reference our plans to engage with the broader community to set up a `social machine' bringing together researchers from various disciplines to help formulate a widely agreed framework for the field.

\section{Comparison with related fields}
\label{sec:comparison}
Turing machines were once conceived as logical automata. We have experienced a `paradigm shift' in the usage of computers, moving away from this narrow characterisation to one according to which machines facilitate a wide range of human interactions. The emergence of `Computer-supported Collaborative Work (CSCW)' \cite{grudin1994computer} is representative of the early phases of this trend. This initial concept has evolved to become the broader field of `Computer-supported collaboration' (CSC).

The Web as a global platform for information access and sharing marked a second essential milestone; in particular, through principles and technologies promoted through Web $2.0$ and the Mobile Web. These developments led to amazing growth in terms of the amounts of content available online and the extent of mass participation. They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopedias, publishing Terabytes of multimedia content, contributing to world-class software, and active participation in defining the agenda for many aspects of our society. This trend towards `prosumerism' is finding more and more adopters in the public and private sector. Governments and enterprises are not only becoming active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organisational management, product development, service offerings, and policy formulation. In this context, a number of terms are used to refer to the ways people interact with each other and with applications: `wisdom of the crowds', `collective intelligence', `open innovation', `crowdsourcing', `human computation', and `social computing'. As we argue below, these terms are related, but not synonymous with `social machines' (see also Figure \ref{fig:socialmachine}).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.5\textwidth]{img/socialmachinesscope_new.png}
\caption{Social machines and related areas} \label{fig:socialmachine}
\end{center}
\end{figure}

Wisdom of the crowds \cite{surowiecki2005wisdom} refers to a principle for decision making that takes into account the information and opinions of a group of people rather than individuals; the use of specific technologies, most notably Web $2.0$, has made it possible for such processes to be carried out at scales inconceivable in the past, and to involve highly diverse and geographically distributed participants. A similar concept, though broader scoped, is collective intelligence, defined in \cite{malone2009harnessing} as `groups of individuals doing things collectively that seem intelligent.' The main difference to what we refer to as `social machines' is with regard to the extent and role of automation. In a social machine, human and computational intelligence coalesce in order to achieve a given purpose. While wisdom of the crowds and collective intelligence place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realisation of hybrid systems in which the two types of components co-exist. As such, theoretical and empirical insights from these two related areas are useful to understand the dynamics of the social structures underlying a social machine, but they are definitely not the only ingredient needed to build operational systems. The question of how human and automatic services can be brought together to achieve optimal results, as well as the actual engineering by which a system is developed, tested, and updated are equally important.

One of the direct consequences of the popularity of the wisdom of the crowds idea was a stronger investment worldwide in open innovation, which can be seen as an application of the concept to business environments, or, in the words of the authors, as a `a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology' \cite{chesbrough2003}. At a more general level, this kind of creativity could be leveraged in almost any domain which benefits from diversity, whilst at the same time ensuring in-time access to a potentially infinite pool of skills and resources not available before the advent of social computing technologies. The term crowdsourcing is typically associated with this larger collection of situations, in which `a job traditionally performed by a designated agent [...] [is outsourced] to an undefined, generally large group of people in the form of an open call' \cite{howe2006crowdsourcing}. Human computation applies human processing power to tackle technical tasks that computers (still) find challenging \cite{von2009human}, typically in areas such as visual, audio, and natural language understanding. These kinds of task are an important part of today's crowdsourcing landscape, in particular on so-called `microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work that take in the range of seconds to minutes to complete. The key difference between social machines and open innovation and crowdsourcing is at the level of the interaction between the social and the machine-driven processing components. The novelty these two other approaches bring in lies in their use of a much larger pool of human resources than traditional work environments \cite{quinn2011human}. There are clear socio-economic implications that their adopters need to deal with in order to optimally make use of this wealth of resources; technology may be needed to assist specific aspects of crowdsourcing projects, from evaluating and rewarding the results produced by the crowd to consolidating and aggregating them into a complete solution. These aspects are equally important for social machines, which, however, also look into the principled combination of human and computational capabilities and the technical means to support them. In comparison to human computation, social machines cover a wider range of scenarios. Human computation is AI-centric and uses people to perform tasks that computers are not (yet) able to tackle (in terms of accuracy) \cite{quinn2011human}; by comparison, we see many successful examples of social machines in which the role of machines is rather to facilitate interactions within groups of people or communities of interest.

An analogous line of reasoning illustrates the overlap between social machines and social computing \cite{parameswaran2007research}. The latter is an area of computer science that refers to systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organisations, and markets' \cite{parameswaran2007research}. As such, compared to the general concept of `Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort. This distinction is even stronger in the case of social machines, which regard the social and the technical components as equal and necessary partners, and study the ways they could be best combined to master the challenges of future socio-technical systems.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{img/polyarchy.pdf}
\caption{A polyarchy of social machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage} \label{polyarchy}
\end{center}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.5\textwidth]{img/dendrogram-both.png}
\caption{Dendrograms of constructs, in blue, and elements (social machines), in red, from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
\end{center}
\end{figure}

>>>>

\numberofauthors{1} %  in this sample file, there are a *total*
% of EIGHT authors. SIX appear on the 'first-page' (for formatting
% reasons) and the remaining two appear in the \additionalauthors section.
%
\author{
% You can go ahead and credit any number of authors here,
% e.g. one 'row of three' or two rows (consisting of one row of three
% and a second row of one, two or three).
%
% The command \alignauthor (no curly braces needed) should
% precede each author name, affiliation/snail-mail address and
% e-mail address. Additionally, tag each line of
% affiliation/address with \affaddr, and tag the
% e-mail address with \email.
%
% 1st. author
\alignauthor
Nigel R. Shadbolt, Daniel A. Smith, Elena Simperl, Max Van Kleek, Yang Yang, \\ Wendy Hall\\
       \affaddr{Web and Internet Science Group}\\
       \affaddr{University of Southampton, UK}\\
       \email{\{nrs,ds,es6g12,emax,yy1402,wh\}@ecs.soton.ac.uk}
}
% There's nothing stopping you putting the seventh, eighth, etc.
% author on the opening page (as the 'third row') but we ask,
% for aesthetic reasons that you place these 'additional authors'
% in the \additional authors block, viz.
%\additionalauthors{Additional authors: John Smith (The Th{\o}rv{\"a}ld Group,
%email: {\texttt{jsmith@affiliation.org}}) and Julius P.~Kumquat
%(The Kumquat Consortium, email: {\texttt{jpkumquat@consortium.net}}).}
%\date{30 July 1999}
% Just remember to make sure that the TOTAL number of authors
% is the number that will appear on the first page PLUS the
% number that will appear in the \additionalauthors section.

@@

The remaining sections are organised as follows. In order to pin down what constitutes a social machine, and establish the boundaries of this emerging field of research, we need to first understand the relationship between social machines and related topics such as human computation, collective intelligence, and crowdsourcing; Section \ref{sec:comparison} provides a summary of this comparative analysis. In Section \ref{sec:framework} we present the framework in terms of the methodology used to devise it, the main clusters of constructs used for classification, and an example of using them to compare a selection of popular social machines via repertory grid analysis. We conclude with our proposed future work, including details concerning the evaluation of the framework. Finally we reference our plans to engage with the broader community to set up a `social machine' bringing together researchers from various disciplines to help formulate a widely agreed classification framework for the field.

\section{Comparison with related fields}
\label{sec:comparison}
Turing machines were once conceived as logical automata. We have experienced a `paradigm shift' in the usage of computers, moving away from this narrow characterisation to one where machines facilitate a wide range of human interactions. The emergence of `Computer-supported Collaborative Work (CSCW)' \cite{grudin1994computer} is representative of the early phases of this trend. This initial concept has evolved to become the broader field of `Computer-supported collaboration' (CSC).

The Web as a global platform for information access and sharing marked a second essential milestone; in particular, through principles and technologies promoted through Web $2.0$ and the Mobile Web. These developments led to amazing growth in terms of the amounts of content available online and the extent of mass participation. They are responsible for hundreds of millions of users all over the globe creating high-quality encyclopaedias, publishing Terabytes of multimedia content, contributing to world-class software, and active participation in defining the agenda for many aspects of our society. This trend towards `prosumerism' is finding more and more adopters in the public and private sector. Governments and enterprises are not only becoming active in open initiatives, but encourage the participation of their customers and employees in taking decisions related to organisational management, product development, service offerings, and policy formulation. In this context, a number of terms are used to refer to the ways people interact with each other and with applications: `wisdom of the crowds', `collective intelligence', `open innovation', `crowdsourcing', `human computation', and `social computing'. As we argue below they are related, but not synonymous with `social machines' (see also Figure \ref{fig:socialmachine}).

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.5\textwidth]{img/Socialmachine1}
\caption{Social machines and related areas} \label{fig:socialmachine}
\end{center}
\end{figure}

Wisdom of the crowds \cite{surowiecki2005wisdom} refers to a principle for decision making that takes into account the information and opinions of a group of people rather than individuals; the use of specific technologies, most notably Web $2.0$, has made it possible for such processes to be carried out at scales inconceivable in the past, and to involve highly diverse and geographically distributed participants. A similar concept, though broader scoped, is collective intelligence, defined in \cite{malone2009harnessing} as `groups of individuals doing things collectively that seem intelligent.' The main difference to what we refer to as `social machines' is with regard to the extent and role of automation. In a social machine, human and computational intelligence coalesce in order to achieve a given purpose. Wisdom of the crowds (WoC) and collective intelligence (CI) clearly place their focus on identifying the situations in which groups of people perform better than individuals, social machines are concerned with the study and realisation of hybrid systems where the two types of components co-exist. As such, theoretical and empirical insights from WoC and CI are useful to understand the dynamics of the social structures underlying such a system, but they are definitely not the only ingredient needed to build operational social machines. The question of how human and automatic services can be brought together to achieve optimal results, as well as the actual engineering by which a system is developed, tested, and updated are equally important.

One of the direct consequences of the popularity of the wisdom of the crowds idea was a stronger investment worldwide in open innovation, which can be seen as an application of the concept to business environments, or, in the words of the authors, as a `a paradigm that assumes that firms can and should use external ideas as well as internal ideas, and internal and external paths to market, as the firms look to advance their technology' \cite{chesbrough2003}. At a more general level, this kind of creativity could be leveraged in almost any domain which benefits from diversity, whilst at the same time ensuring in-time access to a potentially infinite pool of skills and resources not available before the advent of social computing technologies. The term crowdsourcing is typically associated with this larger collection of situations, in which `a job traditionally performed by a designated agent [...] [is outsourced] to an undefined, generally large group of people in the form of an open call' \cite{howe2006crowdsourcing}. Human computation applies human processing power to tackle technical tasks that computers (still) find challenging \cite{von2009human}, typically in areas such as visual, audio, and natural language understanding. These sort of tasks are an important part of today's crowdsourcing landscape, in particular on so-called `microtask' platforms such as Amazon's Mechanical Turk\footnote{\url{https://www.mturk.com/}} or CrowdFlower\footnote{\url{http://crowdflower.com/}}, which offer small financial rewards to an anonymous crowd engaged with atomic units of work that take in the range of seconds to minutes to complete. The key difference between social machines and open innovation and crowdsourcing is at the level of the interaction between the social and the machine-driven processing components. The novelty these two other approaches bring in lies in their use of a much larger pool of human resources then traditional work environments \cite{quinn2011human}. There are clear socio-economic implications that their adopters need to deal with in order to optimally make use of this wealth of resources; technology may be needed to assist specific aspects of crowdsourcing projects, from evaluating and rewarding the results produced by the crowd to consolidating and aggregating them into a complete solution. These aspects are equally important for social machines, which, however, also look into the principled combination of human and computational capabilities and the technical means to support them. In comparison to human computation, social machines cover a wider range of scenarios. Human computation is AI-centric and uses people to perform tasks that computers are not (yet) able to tackle (in terms of accuracy) \cite{quinn2011human}; by comparison, we see many successful examples of social machines in which the role of machines is rather to facilitate interactions within groups of people or communities of interest.

An analogous line of reasoning illustrates the overlap between social machines and social computing \cite{parameswaran2007research}. The latter is an area of computer science which refers to systems that support `the gathering, representation, processing, use, and dissemination of information that is distributed across social collectivities such as teams, communities, organisations, and markets' \cite{parameswaran2007research}. As such, compared to the general concept of `Computer-supported collaboration', social computing puts a greater emphasis on the information management capabilities of groups and communities, and less on the way these capabilities emerge as a joint effort. This distinction is even stronger in the case of social machines, which regard the social and the technical components as equal and necessary partners, and study the ways they could be best combined to master the challenges of future socio-technical systems.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.4\textwidth]{img/polyarchy}
\caption{A polyarchy of social machines, illustrating the infrastructure and frameworks used by social machines, and machine-machine usage} \label{polyarchy}
\end{center}
\end{figure}

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.5\textwidth]{img/dendrogram-both}
\caption{Dendrograms of constructs, in blue, and elements (social machines), in red, from a repertory grid exercise of the top $20$ (ranked by Alexa) social machines from our element set, against our consolidated constructs.} \label{dendrogram}
\end{center}
\end{figure}
